### 2018-05-13 Stochastic Convolutional Sparse Coding/CSCmodel.tex
update draft

\section{Convolutional Sparse Coding (CSC)}
Specifically, the CSC problem is trying to minimize
\begin{equation}
\begin{split}
    \minimize{\filter,\code} & \frac{1}{2}\|\signal - \sum_{k=1}^{K} \filter_k * \code_k \|_2^2 + \lambda \sum_{k=1}^{K}\| \code_k \|_1 \\
    \subjectto & ~ \|\filter_k\|^2_2 \leq 1 ~~ \forall k \in {1,...K},
\end{split}
\end{equation}
where $\signal \in \mathbb{R}^D$ is a D-dimensional signal or a vectorized image~\footnote{In this manuscript, we work on 2D images.}, $\filter_k \in \mathbb{R}^M$ is the $k$-th dictionary and $\code_k$ is the sparse code associated with that dictionary, and it has the same dimension size as $\signal$. $K$ is the number of dictionary filters, and $*$ is the convolution operator. The above equations will be applied to all the training images.

The modern approaches exploit Parseval's theorem and introduce two slack variables to separate the non-smooth $L_1$ penalty term and the $L_2$ constraints, making it feasible to be efficiently computed in frequency domain, and to apply splitting strategy to formulate the problem into ADMM framework, where the two subproblems (updating $\code$ and updating $\filter$) are jointly solved by coordinate descent~\cite{bristow2013fast,heide2015fast,wohlberg2016efficient}. There exists several common issues regarding to the model and the frequency solvers.

\begin{itemize}
  \item Though CSC overcomes the independence assumption held in patch-based learning algorithms, far more variables ($K$ times more), as a payback, will be introduced to represent a single image. This brings in more severe memory and computational burden.
  \item The reconstructed sparse codes show that in the case of $K=100$, the majority of the entries (more than $99.5\%$) do not provide valid information to the represented image, which indicates the subproblem for updating $\code$ solves a highly sparse LASSO problem. Transforming the problem into frequency domain imposes restriction on the use of optimization strategies for this very specific problem.
  \item The prior work shows its efficiency of solving the CSC problem in frequency domain, while this is only applicable for updating $\code$, and it does not hold for updating $\filter$. The dictionary filters usually have much smaller spatial support than the dimension size of the sparse codes ($M \ll D$), while in order to tackle the problem in frequency domain, it requires to process the $\filter$-subproblem in the same dimensions as the sparse codes, and then project the results onto its small spatial support.
\end{itemize}