### 2018-05-15 Stochastic Convolutional Sparse Coding/Conclusion.tex

\section{Conclusions}
In this work, we present a novel stochastic subsampling strategy for updating sparse codes and it successfully improves the runtime performance to solve the CSC problem in spatial domain, which applies to both batch mode and online-learning mode, as well as generating effective and proven outcomes. Since the subproblem of updating the codes is a highly sparse Lasso, specific optimization strategies can be applied to further accelerate the computation, for instance the idea proposed in, which uses coordinate descent and skip unnecessary updates by the method of safe screening. Notice that transforming the problem into Fourier domain cannot benefit from these kinds of speedup strategies. In addition, we show the capability of the developed online algorithm to learn representative over-complete dictionary from large datasets, and the availability of the dictionary is further verified by the application of image inpainting. It can be foreseen that this capability has widespread applications in the audio and image related tasks, and higher dimensional signal processing.