### 2018-11-08 cvpr2019/Conclusion.tex

\section{Conclusions}
In this work, we present a novel stochastic subsampling strategy for
solving the CSC problem in the spatial domain. This method
significantly improves the runtime performance over the prior
frequency-domain solvers, which applies to both batch mode and
online-learning mode.  The proposed algorithm, for the first time,
demonstrates the feasibility that tackling the CSC problem in spatial
domain while still holding, or even improving the runtime
efficiency. Since the subproblem of updating the code is a highly
sparse LASSO, other specific optimization strategies can be applied to
further accelerate the computation, for instance the idea proposed
in~\cite{johnson2017stingycd}, which solves the LASSO problem by
coordinate descent and skips unnecessary updates using the method of
safe screening~\cite{ghaoui2012Swfe}. It is worth emphasizing that
Fourier-domain methods cannot benefit from these kinds of speedup
strategies.

In the future, alternative subsampling strategies can be studied to
improve learning efficiency. Instead of employing uniform subsampling
strategy, for instance one could employ an importance or correlation
based sampling strategy that considers the inter-dependency between the
signals and the sparse codes.

We have also shown the capability of the developed online algorithm to
learn representative and meaningful over-complete dictionary from
arbitrary large datasets, and the availability of the dictionary is
further verified by the application of image inpainting. It can be
foreseen that this capability has widespread applications in audio
and image related tasks, and higher dimensional signal processing. The
source code for SBCSC and SOCSC is attached in the supplement for
reference.
\note{Do you really mean source code (as in full matlab code), or just
  the pseudo-code?}

% --- DO NOT DELETE ---
% Local Variables:
% mode: latex
% mode: flyspell
% mode: TeX-PDF
% End: